### 2022-05-27 tex/prelim.tex
[ more ] thesis writing

>>>>

\begin{definition}
  A \emph{category} is a collection $\obj$ of objects, a family
  $\hom : \obj \times \obj \to \Set$, for each object $A$ an element
  $\id : \hom(A,A)$, and for each triple of objects $A, B, C$ a function
  ${\circ} : \hom(B,C) \times \hom(A,B) \to \hom(A,C)$ such that the following
  laws hold for any $f : \hom(A,B), g : \hom(B,C), h : \hom(C,D)$.
  \begin{itemize}
    \item $\id \circ f = f$
    \item $f \circ \id = f$
    \item $(h \circ g) \circ f = h \circ (g \circ f)$
  \end{itemize}
  I sometimes write $f; g$ for $g \circ f$.
\end{definition}

I give a somewhat unusual definition of \emph{monoidal categories} via
\emph{multicategories}.
I use this approach to capture the purpose of monoidal categories in many
examples, particularly in algebra and programming languages: to provide a way to
encode morphisms of arbitrary arity as unary morphisms by representing $n$-ary
domains as internal $n$-ary products.

I choose to present multicategories in a style where composition models
simultaneous substitution rather than single substitution.
This choice is consistent with the rest of this thesis, which concerns
simultaneous substitution.

\begin{definition}
  A \emph{multicategory} is a collection $\obj$ of objects, a family
  $\hom : \List \obj \times \obj \to \Set$ (with $\hom([A_1, \ldots, A_n], B)$
  written $\hom(A_1, \ldots, A_n; B)$), for each object $A$ an element
  $\id : \hom(A; A)$, and for each list of lists $\Psi$, each list $\Gamma$, and
  each object $A$, a function
  ${\circ} : \hom(\Gamma; A) \times \plr{\prod_i \hom(\Psi_i; \Gamma_i)} \to
  \hom\plr{\sum_i \Psi_i; A}$ such that the following laws hold.
  \begin{itemize}
    \item $\id \circ f = f$
    \item $f \circ \plr{\lambda i.~\id} = f$
    \item $h \circ \plr{\lambda i.~g_i \circ \lambda j.~f_{ij}} =
      (h \circ \lambda i.~g_i) \circ \lambda ij.~f_{ij}$
  \end{itemize}
\end{definition}

\begin{example}
  The following give rise to multicategories:
  \begin{itemize}
    \item Sets and multi-ary functions (functions of the form
      $A_1 \times \cdots \times A_n \to B$)
    \item $K$-vector spaces and multilinear maps
    \item For any simple type theory with sequents of the form
      $\Gamma \vdash A$, where $\Gamma$ is a list of types and $A$ is a type,
      types and open terms (up to $\beta\eta$-equivalence) form a multicategory.
  \end{itemize}
\end{example}

Related to the multicategory of $K$-vector spaces and multilinear maps is the
multicategory of commutative monoids and multilinear maps.
I use the latter later in this chapter\todo{forward reference}, so I will
develop it fully as an example.

\begin{example}
  The multicategory of \emph{commutative monoids and multilinear maps} has as
  objects commutative monoids.
  A morphism of type $A_1, \ldots, A_n \to B$ is a function
  $f : A_1 \times \cdots \times A_n \to B$ such that the following laws hold:
  \begin{itemize}
    \item $f(x_1, \ldots, 0_{A_i}, \ldots, x_n) = 0_B$
    \item $f(x_1, \ldots, y +_{A_i} z, \ldots, x_n) =
      f(x_1, \ldots, y, \ldots, x_n) +_B f(x_1, \ldots, z, \ldots, x_n)$
  \end{itemize}
  The identity function of type $A \to A$ serves as the identity
  (multi)morphism.
  Composition is given by multi-place composition of functions, which satisfies
  the above laws because in a composite $g \circ [f_1, \ldots, f_n]$, a
  commutative monoid operation in the domain of the composite bubbles through
  one of the $f_i$ to become one operation in position $i$ of the domain of $g$,
  and then bubbles through $g$.
  That this identity and composition follow the equational laws follows
  straightforwardly by considering the underlying functions.
\end{example}

\begin{definition}
  A multicategory $\C$ gives rise to its \emph{category of contexts} $\C^*$ as
  follows:
  \begin{itemize}
    \item $\obj(\C^*) \coloneqq \List(\obj\C)$
    \item $\C^*\plr{\sum_i \Psi_i, \Delta} \coloneqq \prod_i \C(\Psi_i; \Delta_i)$
      for any $\Psi : \List(\List(\obj\C))$
    \item $\id^* : \Gamma \to^* \Gamma \coloneqq \overline\id$
    \item $(\sigma : \Delta \to^* \Theta) \circ^* (\rho : \Gamma \to^* \Delta) :
      (\Gamma \to^* \Theta) \coloneqq \lambda i.~\sigma_i \circ \lambda j.~\rho_{ij}$
      \todo{Still wrong}
  \end{itemize}
\end{definition}

\begin{definition}
  Each multicategory $\C$ \emph{restricts} to a category $\C^1$, where
  $\C^1(A, B) \coloneqq \C(A; B)$ and identity and composition are as in $\C$.
\end{definition}

\begin{definition}\label{def:tensor-product}
  A \emph{tensor product} in a multicategory $\C$ of objects $A$ and $B$ is an
  object $A \otimes B$ and a map ${\otimes} : A, B \to A \otimes B$ such that
  any morphism $f : \Gamma, A, B, \Delta \to C$ uniquely factors as:

  \begin{tikzcd}
    \Gamma, A, B, \Delta \arrow[d,"f"']
    \arrow[r,"{\overline\id, \otimes, \overline\id}"]
    & \Gamma, A \otimes B, \Delta
    \arrow[dl,dashed,"f^{\Gamma, A \otimes B, \Delta}"] \\
    C
  \end{tikzcd}

  When it is clear from context which two objects are being multiplied, I will
  write just $f^{\otimes}$ for the unique morphism.
\end{definition}

\begin{definition}\label{def:tensor-unit}
  A \emph{tensor unit} in a multicategory $\C$ is an object $I$ and a map
  $I : {\cdot} \to I$ such that any morphism
  $f : \Gamma, \Delta \to C$ uniquely factors as:

  \begin{tikzcd}
    \Gamma, \Delta \arrow[d,"f"']
    \arrow[r,"{\overline\id, I, \overline\id}"]
    & \Gamma, I, \Delta \arrow[dl,dashed,"f^{\Gamma, I, \Delta}"] \\
    C
  \end{tikzcd}

  When it is clear from context where the $I$ is being introduced, I will
  write just $f^I$ for the unique morphism.
\end{definition}

\begin{lemma}\label{thm:tensor-product-terms}
  For each morphism $f : \Gamma \to A$ and $g : \Delta \to B$, we have a
  morphism $(f, g) : \Gamma, \Delta \to A \otimes B$ such that we have, for any
  $h : \Gamma', A, B, \Delta' \to C$, that
  $h^{\otimes} \circ [\overline\id, (f, g), \overline\id] =
  h \circ [\overline\id, f, g, \overline\id]$, and that
  $\id = (\id, \id)^{\otimes} : A \otimes B \to A \otimes B$.
\end{lemma}
\begin{proof}
  The composite $(\otimes) \circ [f, g]$ has the required type.

  To prove the first equation, note that
  \begin{align*}
    h^{\otimes} \circ [\overline\id, (f, g), \overline\id]
    &= h^{\otimes} \circ [\overline\id, {\otimes} \circ [f, g], \overline\id] \\
    &= h^{\otimes} \circ [\overline\id, {\otimes}, \overline\id]
    \circ [\overline\id, f, g, \overline\id].
  \end{align*}
  Then, we have that
  $h^{\otimes} \circ [\overline\id, {\otimes}, \overline\id] = h$
  by the defining diagram of the tensor product (\cref{def:tensor-product}).

  To prove the second equation, note that the diagram below commutes.
  Because this is an instance of the diagram in \cref{def:tensor-product},
  uniqueness gives us the desired equation.

  \begin{tikzcd}
    A, B \arrow[d,"{\otimes}"'] \arrow[r,"{\otimes}"]
    & A \otimes B \arrow[dl,"\id"] \\
    A \otimes B
  \end{tikzcd}
\end{proof}

\begin{lemma}\label{thm:tensor-unit-terms}
  For any $h : \Gamma', \Delta' \to C$, we have that
  $h^I \circ [\overline\id, I, \overline\id] =
  h \circ [\overline\id, \overline\id]$
  and $\id = I^I : I \to I$.
\end{lemma}
\begin{proof}
  The first equation is just the commuting diagram of \cref{def:tensor-unit},
  modulo composition with identity morphisms.
  For the second equation, note that the diagram below commutes.
  Because this is an instance of the diagram in \cref{def:tensor-unit},
  uniqueness gives us the desired equation.

  \begin{tikzcd}
    {\cdot} \arrow[d,"I"'] \arrow[r,"I"]
    & I \arrow[dl,"\id"] \\
    I
  \end{tikzcd}
\end{proof}

\begin{lemma}
  In a multicategory $\C$ with all tensor products, $\otimes$ induces a
  bifunctor ${\otimes} : \C, \C \to \C$.
\end{lemma}
\begin{proof}
  The action on objects is given by $\otimes$.
  The action on morphisms $f : A \to A'$ and $g : B \to B'$ is given by
  $(f, g)^{\otimes}$.
  Preservation of identity is already proved in \cref{thm:tensor-product-terms},
  and preservation of composition is left as an exercise.
\end{proof}

\begin{lemma}
  For any tensor product and tensor unit, we have the following natural
  isomorphisms:
  \begin{itemize}
    \item $I \otimes A \cong A$
    \item $A \otimes I \cong A$
    \item $(A \otimes B) \otimes C \cong A \otimes (B \otimes C)$
  \end{itemize}
\end{lemma}

\begin{definition}\label{def:monoidal-category}
  A \emph{monoidal category} is a multicategory equipped with a tensor product
  and tensor unit.
  %A \emph{monoidal category} is a category $\C = (\obj, \hom, \id, \circ)$ with
  %functors $I : 1 \to \C$ and ${\otimes} : \C \times \C \to \C$ together with
  %the following natural isomorphisms:
  %\begin{itemize}
  %  \item $\lambda : (I \otimes {-}) \simeq {-}$
  %  \item $\rho : ({-} \otimes I) \simeq {-}$
  %  \item $\alpha :
  %    (({-} \otimes {-}) \otimes {-}) \simeq ({-} \otimes ({-} \otimes {-}))$
  %\end{itemize}
  %such that the following coherence conditions hold:
  %\begin{itemize}
  %  \item
  %    \begin{tikzcd}
  %      (A \otimes I) \otimes B \arrow[dr,"\rho_A \otimes B"'] \arrow[rr,"\alpha_{A,I,B}"]
  %      && A \otimes (I \otimes B) \arrow[dl,"A \otimes \lambda_B"] \\
  %      & A \otimes B
  %    \end{tikzcd}
  %  \item
  %    \begin{tikzcd}
  %      & (A \otimes B) \otimes (C \otimes D)
  %      \arrow[dr,"\alpha_{A,B,C \otimes D}"] &
  %      \\
  %      ((A \otimes B) \otimes C) \otimes D
  %      \arrow[ur,"\alpha_{A \otimes B,C,D}"]
  %      \arrow[d,"\alpha_{A,B,C} \otimes D"]
  %      &&
  %      A \otimes (B \otimes (C \otimes D))
  %      \\
  %      (A \otimes (B \otimes C)) \otimes D
  %      \arrow[rr,"\alpha_{A,B \otimes C,D}"]
  %      &&
  %      A \otimes ((B \otimes C) \otimes D)
  %      \arrow[u,"A \otimes \alpha_{B,C,D}"]
  %    \end{tikzcd}
  %\end{itemize}
\end{definition}

\begin{remark}
  We can talk about monoidal products on a category $\C$ by finding a
  multicategory which restricts to $\C$ and constructing tensor products in it.
\end{remark}

\begin{definition}
  A \emph{monoid} in a multicategory is an object $M$ and multimorphisms
  $\eta : {\cdot} \to M$ and $\mu : M, M \to M$ such that the following
  equations hold:
  \begin{itemize}
    \item $[\eta, \id]; \mu = \id : M \to M$
    \item $[\id, \eta]; \mu = \id : M \to M$
    \item $[\mu, \id]; \mu = [\id, \mu]; \mu : M, M, M \to M$
  \end{itemize}
\end{definition}

\begin{definition}
  TODO: \emph{symmetric monoidal category}
\end{definition}

\begin{definition}
  TODO: \emph{commutative monoid in a symmetric monoidal category}
\end{definition}

Algebra

\begin{definition}
  A \emph{semiring} is a monoid in the multicategory of commutative monoids and
  multilinear maps.
  Unpacked, this means that we have a set $\mathscr R$ together with elements
  $0$ and $1$ and binary operators $+$ and $\cdot$ (with $\cdot$ usually written
  as juxtaposition) such that the following hold for all
  $x, y, z \in \mathscr R$.
  \begin{itemize}
    \item $0 + x = x$; $x + 0 = x$; $(x + y) + z = x + (y + z)$; $x + y = y + x$
    \item $1x = x$; $x1 = x$; $(xy)z = x(yz)$
    \item $0x = 0$; $x0 = 0$; $(x + y)z = xz + yz$; $x(y + z) = xy + xz$
  \end{itemize}
\end{definition}

### 2022-05-27 tex/semantics.tex
[ more ] thesis writing

Our aim with the following traversals is to derive a recursion scheme for
terms which abstracts over some of the details of variable binding and
changes to usage annotations.
We follow the structure given by \citet{AACMM20} closely, with
the main difference having come already in the syntax and environments.

\subsection{Renameability}

{\color{red}(Maybe this goes in the Environment section.)}
We saw in \cref{sec:env} that our notion of \emph{environment} specialises to
\emph{simultaneous renaming} by setting $\V$ to be $\sqni$.

\ExecuteMetaData[\Renamingtex]{Ren}
\ExecuteMetaData[\Renamingtex]{Boxr}
\ExecuteMetaData[\Renamingtex]{Renameable}

\begin{lemma}[\AgdaFunction{ren\textasciicircum{}Env}]\label{thm:env-ren}
  If $\plr{-} \sdtstile{}\V A$ is renameable for all $A$, then so is
  $\plr{-} \env\V \Delta$ for all $\Delta$.
\end{lemma}
\begin{proof}
  For renameable $\V$, we can implement $\mathrm{lift}$ as in
  \cref{thm:env-comp} with $\U \coloneqq {\sqni}$ and
  $\V \coloneqq \W \coloneqq \V$.
  Renaming the environment then becomes an instance of environment composition.
\end{proof}

\subsection{The Kripke function space}

Earlier, when talking about syntax, we saw \AgdaFunction{Scope} as a simple way
to let an open family admit context extensions.
However, a key component of our generic semantic traversal is to make use of
the open family \AgdaBound{$\V$} of \emph{values}, which are the sort of thing
we store in an environment.
The definition \AgdaFunction{Kripke} gives an alternative to
\AgdaFunction{Scope} which interprets the newly bound variables via a
requirement of $\V$-values rather than extra assumptions for the
$\C$-computation.

@@

\AgdaFunction{Wrap} is a device that turns any type family into an equivalent
type family that is judgementally injective in its indices.
It turns the type family into a parametrised record with a single field
\AgdaField{get} whose type is the type in the body of the $\lambda$-abstraction.
For understanding the meaning of the type family, \AgdaFunction{Wrap} can be
ignored.
If $\Delta$ is of the form $\gr{s_1}B_1, \ldots, \gr{s_n}B_n$, then
\ExecuteMetaData[\Snippetstex]{KripkeVCDGA}\ is equivalent to
\ExecuteMetaData[\Snippetstex]{KripkeExpanded}.
That is to say, the Kripke function is expecting a value for each newly bound
variable, at the multiplicity of its annotation, together with the resources
supporting each of those values.
The need for \AgdaFunction{$\Box^r$} is exemplified by the following construct
\AgdaFunction{reify}, where we weaken \AgdaBound{$\Gamma$} by a $\gr0$ed-out
version of \AgdaBound{$\Delta$}.
The \AgdaBound{$\Delta$} then gets filled in by the $\V$-values.

When \AgdaBound{$\V$} supports identity environments (as per \cref{thm:env-id}),
we can coerce a \AgdaFunction{Kripke} function into just a
\AgdaBound{$\C$}-value in an extended context.
To show this, we instantiate the Kripke function with the renaming
$\swarrow^r : \Gamma, \gr0\delta \env\sqni \Gamma$ to extend the scope, and
pass it an argument environment
$\plr{\id^\V; \searrow^r} : \gr0\gamma, \Delta \env\V \Delta$ to fill in the
extended part.
Post-composition of a renaming onto an environment comes from
\cref{thm:env-postren}.

\ExecuteMetaData[\Syntactictex]{reify}

We tend to use \AgdaFunction{reify} whenever \AgdaBound{$\V$} supports
identity environments.
In this paper, all examples support identity environments.
However, \citet[p.~27]{AACMM20} give some important examples that do not
support identity environments, and thus cannot use \AgdaFunction{reify}.
The feature that causes the lack of support for identity environments is that
a semantics can make use of the fact that only variables of particular kinds
are bound by the syntax.
In the examples of \citeauthor{AACMM20}, a bidirectionally typed language only
binds variables that synthesise their type, as opposed to those whose type is
checked.
The semantics of type-checking and elaboration rely on variables synthesising
their type, so \AgdaBound{$\V$} is chosen to cover only those variables.
Instead of using \AgdaFunction{reify}, we must observe that each syntactic form
only binds such synthesising variables.
Similar phenomena would appear in, say, a call-by-value language where
variables are values (not computations), or a polarised language where
variables always have a polarity matching their type.

\subsection{Semantic traversal}

The aim of this subsection is to give an alternative recursion principle for
terms that incorporates some of the environment-handling seen in the
implementations of renaming and substitution.
The rest of this section assumes the following: a renameable open family
\AgdaBound{$\V$} that embeds into the open family \AgdaBound{$\C$}, and an
algebra for a layer of syntax at \AgdaBound{$\C$}.

@@

For \AgdaFunction{body}, we are given a subterm \AgdaFunction{M}, to which we
want to apply \AgdaFunction{semantics}.
To do so, we need an extended version of the initial environment
\AgdaBound{$\rho$}.
The best we can achieve without identity environments for \AgdaBound{$\V$} is
a Kripke function returning an extended environment.
We use \AgdaFunction{mapK$\C$} to post-compose this function by the
\AgdaSymbol{$\lambda$}-function taking an extended environment and acting with
it on \AgdaBound{M}.

>>>>

Given a $\V$-environment $\Gamma \Rightarrow \Delta$, the function
\AgdaFunction{semantics} we define in this section assigns a
$\C$-value in context $\Gamma$ to every term in context $\Delta$,
where $\C$ is an \AgdaFunction{OpenFam} being the carrier of the
semantic interpretation of terms ($\V$ being the semantic
interpretation of variables). Before we can define
\AgdaFunction{semantics}, we need to treat recursion through rules'
premises (\cref{sec:functorial}) and extension of environments when
going under variable binders (\cref{sec:kripke}).

% Our goal in this section is to define \AgdaFunction{semantics}, a
% recursor that turns a term into a \AgdaBound{$\C$}-value using a
% \AgdaBound{$\V$}-environment, in a type preserving way:\bob{Get rid of
%   ``body'' here}

% \ExecuteMetaData[\Semanticstex]{semantics-type}

% The \AgdaBound{$\V$} and \AgdaBound{$\C$} are \AgdaFunction{OpenFam}s,
% representing the interpretations of variables and terms
% respectively. In \cref{sec:traversal} we will see the data that must
% be provided to make a \AgdaFunction{semantics} for a given
% system. Before that, we must see how to deal with the two complicated
% features of our syntax: the usage annotations (\cref{sec:functorial})
% and variable binding (\cref{sec:kripke}). \todo{fwd ref to where these are used}

\subsection{A layer of syntax is functorial}\label{sec:functorial}

A basic property of the universe of syntaxes we described in \cref{sec:syntax}
is that every syntax supports a functorial action on subterms, realised by the function \AgdaFunction{map-s}.
Its type says that to map a function \AgdaBound{f}
over a layer of syntax, there must be a linear map \AgdaBound{F} relating the
domain and codomain usage contexts, and \AgdaBound{f} should be usable
wherever the domain and codomain usage contexts are similarly related by
\AgdaBound{F}.

\ExecuteMetaData[\Maptex]{map-s-type}

This generality is needed because usage contexts change between
a term and its immediate subterms---they are decomposed according to the bunched connectives used in the rules.
\AgdaBound{X} and \AgdaBound{Y} are \AgdaFunction{ExtOpenFam}s, with
\AgdaBound{$\Theta$} being the context extension for a subterm (i.e., the
variables newly bound in that subterm).
Unlike usage annotations, types in the contexts \AgdaBound{$\gamma$} and \AgdaBound{$\delta$}, and the conclusion types implicit here, are preserved throughout.
This is the essence of the usage annotation based approach---we use traditional techniques for variable binding, with the usage annotations layered on top.

The heart of \AgdaFunction{map-s} is \AgdaFunction{map-p}, which recursively
works through the structure \AgdaBound{ps} of premises of the rule applied,
acting on each subterm it finds.
Here, particularly in the clauses for \AgdaInductiveConstructor{`$\sep$} and
\AgdaInductiveConstructor{`$\cdot$}, we see why it is not enough for the
function on subterms to apply at usage contexts \AgdaBound{P} and \AgdaBound{Q}
--- rather, it also needs to apply at any similarly related \AgdaBound{P$'$}
and \AgdaBound{Q$'$}.
In the case of \AgdaInductiveConstructor{`$\sep$}, we have that
$\grP \leq \grP_M + \grP_N$, with \AgdaBound{M} and \AgdaBound{N} being
collections of subterms in usage contexts $\grP_M$ and $\grP_N$, respectively.
Linearity of \AgdaBound{F} yields $\grQ_M$ and $\grQ_N$ such that
$\grQ \leq \grQ_M + \grQ_N$ and we use \AgdaFunction{map-p} recursively at
$(\grP_M, \grQ_M)$ and $(\grP_N, \grQ_N)$ on \AgdaBound{M} and \AgdaBound{N}.
The cases for \AgdaInductiveConstructor{`$\cdot$} and
\AgdaInductiveConstructor{`$I^*$} are similar, each using a different aspect
of linearity.
In contrast, the cases for \AgdaInductiveConstructor{`$\dot1$} and
\AgdaInductiveConstructor{`$\dot\times$}, which are the only constructors used in fully structural
systems, do not involve any changes in the usage contexts.

\ExecuteMetaData[\Maptex]{map-p}

\subsection{The Kripke function space}\label{sec:kripke}

At this point we introduce a minor generalisation to
\AgdaFunction{OpenFam} and \AgdaFunction{ExtOpenFam}:
\AgdaBound{I}\AgdaSpace{}\AgdaFunction{---OpenFam} and
\AgdaBound{I}\AgdaSpace{}\AgdaFunction{---ExtOpenFam}.  We obtain the
definition of \AgdaBound{I}\AgdaSpace{}\AgdaFunction{---OpenFam} by
replacing the textual occurrence of \AgdaBound{Ty} by the parameter
\AgdaBound{I}.

The definition \AgdaFunction{Kripke}\,$\V$\,$\C$\,$\Delta$ is a kind
of function space that describes a $\C$ value parametrised by
$\Delta$-many additional $\V$s (all correctly typed and usage
annotated). It is used to describe how to go under binders in a
Higher-Order Abstract Syntax style---to go under a binder we must
provide semantic interpretations for all the additional variables:

% When going under binders during a recursion, the context will be extended by some $\Theta$. This means that the current environment must be extended with $\Theta$s-worth of $\V$s

% we need the ability to say that

% Kripke V C is given the extension \Theta

% In \cref{sec:terms}, we defined \AgdaFunction{Scope} to let a
% judgement-indexed family admit context extensions. However, a key
% component of our generic semantic traversal is to make use of the open
% family \AgdaBound{$\V$} of \emph{values}, which are the sort of thing
% we store in an environment.  The definition \AgdaFunction{Kripke}
% gives an alternative to \AgdaFunction{Scope} which interprets the
% newly bound variables via a requirement of $\V$-values rather than
% extra assumptions for the $\C$-computation.

@@

\AgdaFunction{Wrap}
is a device that turns any type family into an equivalent type family
that is judgementally injective in its indices, which helps with
Agda's type inference.
It turns the type family into a parametrised
record with a single field \AgdaField{get} whose type is the type in
the body of the $\lambda$-abstraction.
For understanding the meaning of
\AgdaFunction{Kripke}, \AgdaFunction{Wrap} can be ignored.

If $\Delta$ is of the form $\gr{s_1}B_1, \ldots, \gr{s_n}B_n$, then
\ExecuteMetaData[\Snippetstex]{KripkeVCDGA}\ is equivalent to
\ExecuteMetaData[\Snippetstex]{KripkeExpanded}\ by Currying.  That is
to say, the Kripke function is expecting a value for each newly bound
variable, at the multiplicity of its annotation, together with the
resources supporting each of those values. We use the ``magic wand''
function space here to enforce the invariant that the freshly bound
variables have usage annotations that are added to the existing
variables, not shared with them. The use of the
\AgdaFunction{$\Box^r$} modality ensures that we can still use it in
the presence of additional variables introduced by weakening.

\AgdaFunction{Kripke} is functorial in the \AgdaBound{$\C$} argument,
as witnessed by the \AgdaFunction{mapK$\C$} function, which is essentially
post-composition:

\ExecuteMetaData[\Semanticstex]{mapKC}

% is exemplified by the following construct
% \AgdaFunction{reify}, where we weaken \AgdaBound{$\Gamma$} by a $\gr0$ed-out
% version of \AgdaBound{$\Delta$}.
% The \AgdaBound{$\Delta$} then gets filled in by the $\V$-values.

% \bob{Move this para}
% This means that \AgdaBound{A} in the definition of \AgdaFunction{Kripke} has
% type \AgdaBound{I}, rather than specifically \AgdaBound{Ty}.
% We use this generality later in \AgdaFunction{extend}, setting \AgdaBound{I}
% to \AgdaDatatype{Ctx}.

\subsection{Semantic traversal}\label{sec:traversal}

We can now state the data required to implement a traversal assigning
semantics to terms. For open families $\V$ and $\C$, interpreting
variables and terms respectively, we assume that $\V$ is renameable,
that $\V$ is embeddable in $\C$, and that we have an algebra for a
layer of syntax, where bound variables are handled using the Kripke
function space:

% The aim of this subsection is to give an alternative recursion principle for
% terms that incorporates some of the environment-handling seen in the
% implementations of renaming and substitution.
% The rest of this section assumes the following: a renameable open family
% \AgdaBound{$\V$} that embeds into the open family \AgdaBound{$\C$}, and an
% algebra for a layer of syntax at \AgdaBound{$\C$}.

@@

For \AgdaFunction{body}, we are given a subterm \AgdaBound{M}, to
which we want to apply \AgdaFunction{semantics}.  To do so, we need an
extended version of the initial environment \AgdaBound{$\rho$}. We
express this as the generation of a Kripke function that produces the
extended environment given interpretations of the fresh variables. We
take \AgdaBound{$\rho$}, which is an environment covering
\AgdaBound{$\Delta$}, and \AgdaBound{$\sigma$}, which is an
environment covering \AgdaBound{$\Theta$}, and glue them together
using the inductive rules for generating environments, after having
renamed \AgdaBound{$\rho$} via \cref{thm:env-ren} to make it fit the
new context \AgdaBound{$\Gamma^+$} (intended to be
\ExecuteMetaData[\Snippetstex]{GT}):

% The best we can achieve without identity environments for \AgdaBound{$\V$} is
% a Kripke function returning an extended environment.
To define \AgdaFunction{body}, we use \AgdaFunction{mapK$\C$} to
post-compose the environment extension by the
\AgdaSymbol{$\lambda$}-function taking an extended environment and
acting with it on \AgdaBound{M}.

### 2022-05-27 tex/semirings.tex
[ more ] thesis writing

In \cref{sec:lr}, I define the usage-annotated calculus \name{}, which has
appeared previously in my work with Atkey~\cite{WA20}, and can be seen as a
simply typed version of Atkey's dependently typed calculus QTT~\cite{Atkey18}.
The idea of \name{} is to augment simply typed $\lambda$-calculus with
annotations on free variables, and give enough types to manipulate these
annotations.
I use \cref{sec:lnd} to introduce some notation that allows us to restate the
typing rules of \name{} so as to not mention contexts explicitly.
The first problem I tackle with this new syntax is proving admissibility of
substitution.
I follow the methodology from \cref{sec:kits}, which

\section{A usage-annotated calculus}\label{sec:lr}
\input{lr.tex}
\section{Linear natural deduction}\label{sec:lnd}
\input{lnd.tex}
\section{What are linear renaming and substitution?}\label{sec:lrkits}
\input{lrkits.tex}
\section{Properties of linear environments}\label{sec:lenv}
\input{lenv.tex}
\section{Substitution is admissible in \name{}}\label{sec:lrsub}
\input{lrsub.tex}
\section{Adding recursion to \name{}}\label{sec:rec}
\input{rec.tex}
\section{Addendum: (lack of) partiality}\label{sec:part}
\input{part.tex}

>>>>

In \cref{sec:lr}, I define the usage-annotated calculus \name{}, which has
appeared previously in my work with Atkey~\cite{WA21}, and can be seen as a
simply typed version of Atkey's dependently typed calculus QTT~\cite{Atkey18}.
The idea of \name{} is to augment simply typed $\lambda$-calculus with
annotations on free variables, and give enough types to manipulate these
annotations.
I use \cref{sec:lnd} to introduce some notation that allows us to restate the
typing rules of \name{} so as to not mention contexts explicitly.
The first problem I tackle with this new syntax is proving admissibility of
substitution.
I follow the methodology from \cref{sec:kits}, which

\section{A usage-annotated calculus}\label{sec:lr}
\input{lr.tex}
\section{Bunched connectives}\label{sec:lnd}
\input{lnd.tex}
\section{What are linear renaming and substitution?}\label{sec:lrkits}
\input{lrkits.tex}
\section{Properties of linear environments}\label{sec:lenv}
\input{lenv.tex}
\section{Substitution is admissible in \name{}}\label{sec:lrsub}
\input{lrsub.tex}
\section{Adding recursion to \name{}}\label{sec:rec}
\input{rec.tex}
\section{Addendum: (lack of) partiality}\label{sec:part}
\input{part.tex}
